Technique for \earlywarning\ detection of compact binary coalescence and its implications for multi-messenger astronomy

### 2011-06-03 conclusions.tex
merge
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@912 ccdd05d5-56c1-461a-bf98-85f6876523b4

\editorial{We should drive home the point that our method is as fast as the fft convolution but without any latency at all.}%
We have demonstrated a computationally feasible procedure for the rapid and even advance
detection of gravitational waves emitted during the coalescence of neutron
stars and stellar-mass black holes.  These sources are expected to produce
prompt electromagnetic signals and may be the progenitors of some short hard
gamma-ray bursts.  Rapid alerts to the broader astronomical community may
improve the chances of detecting an electromagnetic counterpart in bands from
X-ray down to radio.  We antipate requiring no more than \numcpus\ modern
%
\editorial{We need to do this calculation using the \flops\ counts and the number of templates.}%
%
computer cores to analyze a four-detector network of gravitational-wave data
for binary neutron stars and stellar mass black holes.  This is within the
current computing capabilities of the \textsc{lsc} Data Grid~\cite{LDG}.

The algorithm we described has no intrinsic latency.  However, there are
fundamental and practical latencies associated with the analysis and detection
procedure. For example, the \LIGO{} detectors, data acquisition is synchronized
to a 1/(16~Hz) cadence introducing an up-front latency of 125~ms.  Data
%
\editorial{This is redundant: 1/(16~Hz) and 125~ms are just reciprocals of each other.}%
%
aggregation from the observatories will travel over various networks, each
capable of high bandwith but perhaps only modest latency.  This could amount to
a similar latency of $\sim$100 ms.  Lastly, unless a realtime infrastructure is
adopted post data acquisition, it is likely that there will be an inherent
latency introduced by such infrastructure.  We have shown a prototype
implementation using \gstlal\ that is capable of $\sim$1 s latency. In our
opinion, significant work would have to be done in order to improve upon this
number. However, it should be considered for third-generation detector design.
For example, a tighter integration of analysis and data acquisition would be
beneficial.

\editorial{The way this is worded, it sounds like a big omission.}%
We have omitted discussion of source sky localization. Localization is known to
be poor for signals of low \SNR~\cite{Fairhurst2009}.
However, one should not immediately dismiss the practical usefulness of a
poorly localized source. Even with poor localization, it should be possible to
begin downselecting what observatories could view a potential signal and for
such observatories to begin any necessary prerequisite activities. In future
works we wil explore more rigorously the pointing prospects with realistic
simulations using the infrastructure and techniques described in this work.
\editorial{nvf: It's a bit pie-in-the-sky, but we could also mention reconfiguring
the signal recycling mirror to optimize SNR at merger. There's a lot of science there.}

>>>>

\editorial{We should drive home the point that our method is as fast as the fft convolution but without any latency at all.}%
We have demonstrated a computationally feasible procedure for the rapid and even advance
detection of gravitational waves emitted during the coalescence of neutron
stars and stellar-mass black holes.  These sources are expected to produce
prompt electromagnetic signals and may be the progenitors of some short hard
gamma-ray bursts.  Rapid alerts to the broader astronomical community may
improve the chances of detecting an electromagnetic counterpart in bands from
X-ray down to radio.  We anticipate requiring $\sim$\numcpus\ modern
%
\editorial{We need to do this calculation using the \flops\ counts and the
number of templates. CHAD: Agreed, but if you do it by the books it seems
misleading.  I have adjusted it to 600.  This number is taken from assuming
that 1 657 template sub-bank can be filtered on one core and dividing the total
number of templates (10$^5$) by 657 which gives you 150 cores per detector. I
put it in section 2. }%
%
computer cores to analyze a four-detector network of gravitational-wave data
for binary neutron stars and stellar mass black holes.  This is within the
current computing capabilities of the \textsc{lsc} Data Grid~\cite{LDG}.

The algorithm we described has no intrinsic latency.  However, there are
fundamental and practical latencies associated with the analysis and detection
procedure. For example, the \LIGO{} detectors, data acquisition is synchronized
to a 1/(16~Hz) cadence introducing an up-front latency.  Data
aggregation from the observatories will travel over various networks, each
capable of high bandwith but perhaps only modest latency.  This could amount to
a similar latency of $\sim$100 ms.  Lastly, unless a realtime infrastructure is
adopted post data acquisition, it is likely that there will be an inherent
latency introduced by such infrastructure.  We have shown a prototype
implementation using \gstlal\ that is capable of $\sim$1 s latency. In our
opinion, significant work would have to be done in order to improve upon this
number. However, it should be considered for third-generation detector design.
For example, a tighter integration of analysis and data acquisition would be
beneficial.

\editorial{The way this is worded, it sounds like a big omission.}%
We have omitted discussion of source localization though point
the reader to some theoretical estimates \SNR~\cite{Fairhurst2009}.
%However, one should not immediately dismiss the practical usefulness of a
%poorly localized source. Even with poor localization, it should be possible to
%begin downselecting what observatories could view a potential signal and for
%such observatories to begin any necessary prerequisite activities. 
In future works we will explore more rigorously the pointing prospects with
realistic simulations using the infrastructure and techniques described here.
\editorial{nvf: It's a bit pie-in-the-sky, but we could also mention reconfiguring
the signal recycling mirror to optimize SNR at merger. There's a lot of science there.}
